The effects of binarity on planet occurrence rates measured by transit surveys

### 2018-02-14 paper/main.tex
improve text (minor)

\shortauthors{Bouma et al.}
\shorttitle{Binary Biases}

@@

Though the details of the planet radius distribution $\Gamma_i(r)$ at
$r<2r_\oplus$ are currently an active topic of research, by making
plausible assumptions we can investigate how neglecting binarity might
bias measurements in this regime.  For instance, consider a broken
power-law:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2r_\oplus \\
        {\rm constant} & \text{for } r\leq2r_\oplus.
    \end{cases}
    \right.
    \label{eq:model3_radius_distribution}
\end{align}
At apparent radii $r\a > 2r_\oplus$, the apparent rate density in
this model is the same as that in
Section~\ref{sub:powerlaw_varying_binaries}.  At smaller apparaent
radii the equations are tedious, but still tractable.
For simplicity, we insert Equation~\ref{eq:model3_radius_distribution}
into Equation~\ref{eq:general_Gamma_a}, and integrate using a computer
program. We refer the interested reader to our online
implementation\footnote{\url{https://github.com/lgbouma/binary_biases}}.
As before, $L\propto M^\alpha \propto R^\alpha$, and $\alpha=3.5$,
$\beta=0$, $\gamma=-2.92$.  The output, shown in
Figure~\ref{fig:occ_rate_model_3_log}, is validated using analytic
predictions in the $r\a > 2r_\oplus$ and the $r\a <
2r_\oplus/\sqrt{2}$ regimes.

@@

\paragraph{Hot Jupiter rate discrepancy}
While unresolved binaries may bias $\eta_\oplus$ measurements, our
approach suggests that they do not influence the hot Jupiter rate
discrepancy.  The ``discrepancy'' is that RV surveys find roughly
twice as many hot Jupiters per Sun-like star as do transit surveys
($\approx 1\%$~vs.~$\approx 0.5\%$, respectively).  We summarize hot
Jupiter rates reported by different surveys in
Table~\ref{tab:hj_rates}.

@@

\paragraph{Fraction of detected planets with binary companions}
A related effect is that some detected planets are more likely to be
in binary systems than others.  Notably,
Figure~\ref{fig:frac_model_3} predicts that the fraction of detected
planets with binary companions should increase by $\approx 6-12\%$
going from $2r_\oplus$ to $1.4r_\oplus$.

@@

\paragraph{Connecting high-resolution imaging to occurrence rates}

@@

Even if the multiplicity of every star that {\it Kepler} surveyed were
known, the problem would not be solved. In this scenario, it would be
possible to tabulate planet occurrence for single and binary star
systems separately.  The number of planets per single star would thus
be known.  However the number of planets per primary and secondary
would still be convolved.  An observational approach to
separating the two populations might be to analyze the centroids, or the transit
durations (and thus host star densities)
for a representative sample of planets in binary systems.

A theory-driven solution might entail probabilistic population
inference, with a parametrized model for planet populations around
primary and secondary stars.  Such an approach might find the
populations to differ, given that secondary stars are less
massive, and so are perhaps more likely to host smaller
planets~\citep{dressing_occurrence_2015}.  Given the difficulty of
writing a correct likelihood function in a survey with systematic
biases, this might only be feasible by forward-modeling the survey in
the framework of approximate Bayesian
computation~\citep[\textit{e.g.},][]{morehead_understanding_2016}.

@@

Applying Equation~\ref{eq:general_Gamma_a}, we then showed that:
\begin{itemize}
%
\item If single, primary, and secondary stars all host the same number
  of planets per star, and the planet radius distribution is a power
    law with the exponent reported by~\citet{howard_planet_2012} at $r
    > 2r_\oplus$, then binarity influences apparent planet occurrence
    rates at the few percent level. This applies for apparent radii
    from $2r_\oplus$ to $17r_\oplus$ (Section~\ref{sec:model_2}).
%
\item Assuming a broken-power law planet radius distribution, with the
  \citet{howard_planet_2012} exponent above $2r_\oplus$ and a constant
    occurrence below $2r_\oplus$, the rate density for radii below
    $2r_\oplus$ is overestimated. The overestimate, $\Delta \Gamma_0 =
    |\Gamma_0 - \Gamma\a|/\Gamma_0$, can easily reach $50\%$
    (Figure~\ref{fig:occ_rate_model_3_log}).  Although this is smaller
    than current systematic uncertainties on the occurrence rates of
    Earth-sized planets, this implies that binarity could eventually
    become an important component of the $\eta_\oplus$ error budget.
%
\item Binarity blurs sharp features in the radius distribution
  (Figure~\ref{fig:model_4}), to a degree that could affect precise
    measurements of the depth, width, and slope of
    the~\citet{fulton_california-_2017} radius gap.
    %in the event that
    %planets are not carefully vetted with high-resolution imaging.
%
\item Binarity does not lead to smaller apparent HJ occurrence rates
  (Figure~\ref{fig:gaussian_HJ}).  This assumes a Gaussian planet
    radius distribution, similar to that reported
    by~\citet{petigura_CKS_2017}.
%
\item Detected planets with $r\a \gtrsim 0.5r_\oplus$ that are
  revealed by high-resolution imaging surveys to exist in binaries are
    more likely to orbit the primary (Figure~\ref{fig:frac_model_3}).
%
\item Near the break in occurrence rate as a function of planet
  radius ($\approx 2r_\oplus$), the fraction of detected planets with
    binary companions should increase by roughly $5$~-~$10\%$
    (Figure~\ref{fig:frac_model_3}).
\end{itemize}

These results are only strictly applicable for idealized transit
surveys meeting the criteria mentioned above.  For real transit
surveys, although our approach is only suggestive, we hope that it
provides a rough estimate for the systematic errors that can be
incurred by ignoring binarity when calculating planetary occurrence
rates.

>>>>

\shortauthors{Bouma et al.}
\shorttitle{Binarity's Biases}

@@

Though the details of the planet radius distribution $\Gamma_i(r)$ at
$r<2r_\oplus$ are currently an active topic of research, by making
plausible assumptions we can investigate how neglecting binarity might
bias measurements in this regime.  For instance, consider a broken
power-law:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2r_\oplus \\
        {\rm constant} & \text{for } r\leq2r_\oplus.
    \end{cases}
    \right.
    \label{eq:model3_radius_distribution}
\end{align}
At apparent radii $r\a > 2r_\oplus$, the apparent rate density in
this model is the same as that in
Section~\ref{sub:powerlaw_varying_binaries}.  At smaller apparaent
radii the equations are tedious, but still tractable.
For simplicity, we insert Equation~\ref{eq:model3_radius_distribution}
into Equation~\ref{eq:general_Gamma_a}, and integrate using a computer
program. We refer the interested reader to our online
implementation\footnote{\url{https://github.com/lgbouma/binary_biases}}.
The output, shown in Figure~\ref{fig:occ_rate_model_3_log}, is
validated using analytic predictions in the $r\a > 2r_\oplus$ and the
$r\a < 2r_\oplus/\sqrt{2}$ regimes.  As before, $L\propto M^\alpha
\propto R^\alpha$, and $\alpha=3.5$, $\beta=0$, $\gamma=-2.92$.

@@

\paragraph{Hot Jupiter rate discrepancy}
While unresolved binaries may bias $\eta_\oplus$ measurements, our
approach suggests that they do not influence the hot Jupiter rate
discrepancy.  The ``discrepancy'' is that RV surveys find roughly
twice as many hot Jupiters per Sun-like star as do transit surveys
(roughly $1\%$~vs.~$0.5\%$, respectively).  We summarize hot
Jupiter rates reported by different surveys in
Table~\ref{tab:hj_rates}.

@@

\paragraph{Fraction of detected planets with binary companions}
A related effect is that some detected planets are more likely to be
in binary systems than others.  Notably, Figure~\ref{fig:frac_model_3}
predicts that the fraction of detected planets with binary companions
should increase by $6$~to~$12\%$ going from $2r_\oplus$ to
$1.4r_\oplus$.

@@

\paragraph{Connection between high-resolution imaging and occurrence rates}

@@

Even if the multiplicity of every star that {\it Kepler} surveyed were
known, the problem would not be solved. In this scenario, it would be
possible to tabulate planet occurrence for single and binary star
systems separately.  The number of planets per single star would
be known.  However the number of planets per primary and secondary
would still be convolved.  An observational approach to separating the
two populations might be to analyze the centroids, or the transit
durations (and thus host star densities) for a representative sample
of planets in binary systems.

A theory-driven solution might entail probabilistic population
inference, with a parametrized model for planet populations around
primary and secondary stars.  Such an approach might be expected to
find that the primary and secondary populations are different, since
single M dwarfs host more small planets than FGK
dwarfs~\citep{dressing_occurrence_2015}.  Given the difficulty of
writing a likelihood function in a survey with systematic
biases, one possible approach might be to forward-model the survey in
the framework of approximate Bayesian
computation~\citep[\textit{e.g.},][]{morehead_understanding_2016}.

@@

Applying Equation~\ref{eq:general_Gamma_a}, we then showed that:
\begin{itemize}
%
\item If single, primary, and secondary stars all host the same number
  of planets per star, and the planet radius distribution is a power
    law with the exponent reported by~\citet{howard_planet_2012} at $r
    > 2r_\oplus$, then binarity influences apparent planet occurrence
    rates at the few percent level. This applies for apparent radii
    from $2r_\oplus$ to $17r_\oplus$ (Section~\ref{sec:model_2}).
%
\item Assuming a broken-power law planet radius distribution, with the
  \citet{howard_planet_2012} exponent above $2r_\oplus$ and a constant
    occurrence below $2r_\oplus$, the rate density for radii below
    $2r_\oplus$ is overestimated. The overestimate, $\Delta \Gamma_0 =
    |\Gamma_0 - \Gamma\a|/\Gamma_0$, can easily reach $50\%$
    (Figure~\ref{fig:occ_rate_model_3_log}).  Although this is smaller
    than current systematic uncertainties on the occurrence rates of
    Earth-sized planets, this implies that binarity could eventually
    become an important component of the $\eta_\oplus$ error budget.
%
\item Binarity blurs sharp features in the radius distribution
  (Figure~\ref{fig:model_4}), to a degree that could affect precise
    measurements of the depth, width, and slope of
    the~\citet{fulton_california-_2017} radius gap.
    %in the event that
    %planets are not carefully vetted with high-resolution imaging.
%
\item Binarity does not lead to smaller apparent HJ occurrence rates
  (Figure~\ref{fig:gaussian_HJ}).  This assumes a Gaussian planet
    radius distribution, similar to that reported
    by~\citet{petigura_CKS_2017}.
%
\item Detected planets with $r\a \gtrsim 0.5r_\oplus$ that are
  revealed by high-resolution imaging surveys to exist in binaries are
    more likely to orbit the primary (Figure~\ref{fig:frac_model_3}).
%
\item Near the break in occurrence as a function of planet
  radius at $2r_\oplus$, the fraction of detected planets with
    binary companions should increase by roughly $5$~to~$10\%$
    (Figure~\ref{fig:frac_model_3}).
\end{itemize}

These results are only strictly applicable for idealized transit
surveys meeting the criteria mentioned above.  For real transit
surveys, although our approach is only suggestive, we hope that it
provides a useful estimate for the systematic errors that can be
incurred by ignoring binarity in calculations of planetary occurrence
rates.
